---
title: "Benchmarking the accuracy of multiple systems estimation for the quantification of modern slavery"
author: "George Lindner, Morris Greenberg, Pierre Gardan and Olivier Binette"
date: 'December 14, 2019'
runtime: shiny
output:
  html_document:
    includes:
      before_body: html/header.html
    code_folding: hide
    df_print: paged
    theme: flatly
    toc: true
    number_sections: false
    toc_depth: 2
    toc_float: 
      collapsed: true
---

```{r, echo=FALSE, message = FALSE}
if(!require(pacman)){install.packages("pacman")}
pacman::p_load("shiny", "shinycssloaders", "knitr", "kableExtra", 
               "tidyverse", "dga", "MCMCpack", "Rcapture", "LCMCR",
               "UpSetR", "DT", "ggplot2", "cowplot", "resample", 
               "pbapply", "parallel")

source("../input/silverman_code/datasets.R")
source("../input/silverman_code/functions.R")
```

# 1. Introduction

Forty million people currently live in slavery, according to the International Labour Organization (ILO) [most recent estimate](https://www.ilo.org/global/publications/books/WCMS_575479/lang--en/index.htm). In other words, about every 1 in 220 person worldwide is trapped in a situation of "forced labour, forced marriage, or under another form of exploitation that a person cannot refuse or leave because of threats, violence, coercion, deception and/or abuse of power". **Modern slavery** is an umbrella term which broadly refers to such forms of slavery that are currently happening. 

The scale of the problem has prompted the United Nations to add, in 2015, the eradication of modern slavery as part of its [Sustainable Development Goals framework](https://sustainabledevelopment.un.org/). This calls for integrated action to end modern slavery and for the development of reliable statistical indicators. 

> "Take immediate and effective measures to eradicate forced labour, end modern slavery and human trafficking and secure the prohibition and elimination of the worst forms of child labour, including recruitment and use of child soldiers, and by 2025 end child labour in all its forms." (*Target 8.7 of UN's Sustainable Development goals.*)


<!--
However, statistical methodology to estimate the number of undetected victims of modern slavery is still in development. Victims of modern slavery are often hidden, making reliable estimates of the number of victims particularly challenging to develop.
-->

## 1.1. The case of the UK

The United Kingdom's quantification of modern slavery, together with their 2015 Modern Slavery Act, has drawn considerable attention to such efforts. [Silverman (2014)](https://www.gov.uk/government/publications/modern-slavery-an-application-of-multiple-systems-estimation); [Bales et al. (2015)](https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2015.00824.x) estimated between 10,000 to 13,000 potential victims using a method called **Multiple Systems Estimation** (MSE). The method aggregates data from multiple sources, such as the police and non-governmental organizations, and estimates the proportion of victims that have not been seen by any of them.

However, follow-up research has shown high sensitivity of the estimates to modelling assumptions, with the Walk Free Foundation providing an [about 10-fold larger estimate](https://downloads.globalslaveryindex.org/ephemeral/GSI-2018_FNL_190828_CO_DIGITAL_P-1576254958.pdf) of about 136 000 victims and a recent analysis [(Whitehead et al., 2019)](https://www.tandfonline.com/doi/abs/10.1080/23322705.2019.1660952?journalCode=uhmt20) even suggesting to abandon attempts at measuring this hidden population. It is unclear if the assumptions underlying standard use of MSE were satisfied in Silverman's study and we are concerned by these developments. MSE methods are used in many sensitive human rights applications and their correct usage is critical to reach truth and maintain credibility. 

## 1.2. Our benchmark

We provide some of the tools necessary to investigate the accuracy and stability of MSE estimates in application to the quantification of modern slavery. Precisely, this interactive document (Section 3) allows to:

1. omit or merge sources of information about potential victims and see the effect on the estimates;
2. estimate the coverage of confidence intervals reported by [Silverman (2014)](https://www.gov.uk/government/publications/modern-slavery-an-application-of-multiple-systems-estimation) through a parametric Bootstrap; and
3. simulate data for a range of realistic model and assess the accuracy of MSE in this context. 

We provide an introduction to MSE in Section 2, the tool is embedded in Section 3 and we sumarize our results in Section 4.

# 2. Multiple systems estimation for the quantification of modern slavery

Multiple systems estimation, also called capture-recapture, is a method which combines multiple sources of information to estimate the size of hidden populations. 
For example, imagine we are trying to estimate the number of fish in a large pond. It is not possible to enumerate all fish, but we are able to take random samples with replacements. A strategy to get an estimate of the unknown total population $N$ is the following:

1. mark a first sample of fish, of size $N_1$, and release them back;
2. take a second sample of fish, of size $N_2$, and count the overlap $m$ with the first sample.

We then have observed a proportion $m/N_1$ of the fish in the first sample. Extrapolating, we assume that our sample of size $N_2$ represents a proportion $m/N_1$ of the total population size $N$. Rearranging the terms, this gives an estimate $\hat{N} = N_1 N_2 / m$.

In the context of modern slavery, the samples are obtained by **organizations** observing **potential victims** and recording the information in **lists**. In Silverman's (2014) study, the data recording observed potential victims is considered in six lists: local authorities (LA), non-government organisations (NG), police forces (PF), government organisations (GO), general public (GP), and national crime agency (NCA).

The overlap between the lists has been computed by a trusted analyst, and the data is summarized below. The rows represent inclusion paterns and the associated count. For instance, the first row, with a "1" under the list LA and with "0" under all other lists, states that exactly 54 potential victims have been observed by local authorities (LA) and by no one else. The row with "1" under both LA and NG, with "0" under other lists, counts individuals recorded by both LA and NG, and by no one else. Using these counts for all possible pattern of overlaps, multiple systems estimation provides an estimate of the total population size, including unobserved  potential victims.

```{r, fig.align="center"}
# HTML rendering of the `UKdat` table.
knitr::kable(UKdat) %>% 
  kable_styling(full_width=FALSE, position="center") %>% 
  scroll_box(height = "200px", width="500px") %>% 
  footnote("The data are considered in six lists, labelled as follows:  LA--Local authorities; NG--Non-government organisations; PF--Police forces; GO--Government organisations; GP--General public; NCA--National Crime Agency.", footnote_as_chunk=T)
```

This data is also plotted below, showing the counts for all observed inclusion patterns.

```{r fig.align="center"}
plot_MSE_counts <- function(data, ...) {
  n = ncol(data) - 1
  dat = uncount(data[, 1:n], data$count)
  upset(dat, nsets=n, ...)
}

plot_MSE_counts(UKdat, 
                sets.bar.color = "cornflowerblue",
                main.bar.color = "cornflowerblue",
                matrix.color = "cornflowerblue",
                mainbar.y.label = "Potential victims count",
                sets.x.label = "")
```
 
## 2.1. Underlying assumptions

In its simplest form, MSE makes a number of assumptions:

1. *Independence:* the lists are independent from one another.

2. *Homogeneity:* on a given list, all the individuals have the same probability of being recorded.

3. *Closed population:* the population size does not change during the study.

4. *Perfect matching:* the overlap between lists is computed without any error.

These are very restrictive and unlikely to be satisfied in reality. Fortunately, we can account for dependence between the lists by also modelling and estimating interaction terms. This is done through the use of a log-linear model with interactions terms.

## 2.2. Log-linear model

Precisely, for every lists subset $A \subset \{\text{LA}, \text{NG}, \text{PF}, \text{GO}, \text{GP}, \text{NCA}\}$, the counts $N_A$ of individuals observed on lists in $A$ and on no other lists are modelled as
$$
  N_A \sim \text{Poisson}\left(\exp\left( \mu + \sum_{i \in A} \alpha_i + \sum_{i,j\in A,\,i\not = j} \lambda_{i,j} \right)\right).
$$
The parameters $\alpha_i$ are main effects and the $\lambda_{i,j}$ are interaction terms. In particular, the number of *unobserved victims* which we are interested about is
$$
  N_\emptyset \sim \text{Poisson}(\exp(\mu)), \quad \mathbb{E}[N_\emptyset] = \exp(\mu).
$$

This log-linear Poisson count model is an approximation to a multinomial model and the parameters can be interpreted in term of joint list inclusion probabilities, although this level of detail is not necessary for our purpose.

## 2.3. Silverman's (2014) `MSEfit` function

As the basis of UK's Home Office 10,000 to 13,000 potential victims estimate, Silverman (2014) merged together the NCA and PF lists, and fitted the Poisson log-linear model, selecting list interaction terms to be including through stepwise AIC model selection. They reported the confidence interval corresponding to the model fit with the selected interaction terms.

This procedure is implemented in the `MSEfit` function provided by [Silverman (2019)](https://arxiv.org/abs/1902.06078).

```{r}
MSEfit(UKdat_5)$CI
```
 
The **abundance** represents the estimated count of potential victims, and the two other numbers are respectively the lower and upper confidence limits.

# 3. Benchmark {.tabset}

There are a number of potential issues with Silverman's `MSEfit` method. Some of these have been investigated in Silverman (2019) and Whitehead et al. (2019), and others we also consider in our benchmark.

First, Silverman (2019) has shown some instabilities of the estimates to operations such as merging and omitting lists. The **data manipulation** section of our benchmark allows the user to merge and omit lists in any way, seeing the effect on the patterns of inclusion and on the MSE estimates.

Second, we study the coverage of Silverman's `MSEfit` method in a parametric bootstrap. For three different datasets, the `UKdat` data with all six lists, the `UKdat_5` data merging the PF and NCA lists, and the `UKdat_4` data obtained by merging PF and NCA, and omitting the GP list. We limited the number of datasets to allow us to precompute the simulations. Otherwise, since the `MSEfit` function is quite slow, the simulations can take up to many hours to run.

Third, we consider simulated data where the user is allowed to modify the true population size, the lists inclusion probabilities, the probability of an individual being "scooped" by one of the list and not found by any other, and the probability of an individual being referred by one of the list to another. The *Inclusion Probability* parameters allows the user to experiment with non-homogeneous list inclusion probabilities.

**Usage recommentation:** <span style="color:red;"> Please allow one minute for the first plot to appear, before begining interacting with the application.</span>


## Data manipulation

Create up to six sources of information by selecting lists to use, or multiple lists to merge together. Fields can be left empty to omit certain sources.

The data obtained for the selection of lists is plotted below, and the results of the `MSEfit` function, with the estimated population size as `abundance` and lower and upper confidence limits, is printed underneath the plot.

```{r}
shinyAppFile("data_application/shiny.R", options = list(
    width = "100%", height = 600
  ))
```

## Bootstrap simulation

Select one of three datasets and a number of iterations for a parametric bootstrap study. An histogram representing the bootstrap estimates and the true population size (full vertical line) shows the accuracy of the `MSEfit` method. The estimated coverage of `MSEfit` on the selected dataset is printed under the histogram.

We recommend using pre-processed data, as otherwise the simulations can take multiple hours to run. Executing this application on a machine with many cores will improve performance.

```{r}
shinyAppFile("simulation/shiny2.R", options = list(
    width = "100%", height = 550
  ))
```

## Generative models



```{r}
shinyAppFile("generative/shiny.R", options = list(
    width = "100%", height = 1400
  ))
```

# 4. Conclusions



