---
title: "Benchmarking the accuracy of multiple systems estimation for the quantification of modern slavery"
author: "The Pythonistas"
date: '2019-12-09'
runtime: shiny
output:
  html_document:
    includes:
      before_body: header.html
    code_folding: hide
    df_print: paged
    theme: flatly
    toc: true
    number_sections: false
    toc_depth: 2
    toc_float: 
      collapsed: true
---


# 1. Introduction

Forty million people currently live in slavery, according to the International Labour Organization (ILO) most recent estimate [ref]. In other words, every 1 in 220 people worldwide is trapped in a situation of "forced labour, forced marriage, or under another form of exploitation that a person cannot refuse or leave because of threats, violence, coercion, deception and/or abuse of power" (ILO definition). The term **modern slavery** broadly refers to all such forms of slavery that are currently happening.

The scale of the problem has prompted the United Nations to add, in 2015, the eradication of modern slavery as part of its [Sustainable Development Goals framework](https://sustainabledevelopment.un.org/):

> Target 8.7: Take immediate and effective measures to eradicate forced labour, end modern slavery and human trafficking and secure the prohibition and elimination of the worst forms of child labour, including recruitment and use of child soldiers, and by 2025 end child labour in all its forms. - United Nations' Sustainable Development Goals

This target calls for integrated action to end modern slavery and for the development of reliable statistical indicators. However, statistical methodology to estimate the number of undetected victims of modern slavery is still in development. Obtaining estimates in more economically developped countries is particularly challenging from a statistical point of view, even though more ressources may be allocated towards this.

## 1.1 The case of the UK

The United Kingdom has been a leader in the fight against modern slavery, with Theresa May stating in 2016:

> "This is the great human rights issue of our time, and as Prime Minister I am determined that we will make it a national and international mission to rid our world of this barbaric evil."


# Tabs example {.tabset}

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions}

shinyUI(sidebarLayout(
  sidebarPanel( 
    
    selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
    
    sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
    
),
  mainPanel(
    plotOutput("plot_output")
  )
))

output$plot_output = renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.

