---
title: "Benchmarking the accuracy of multiple systems estimation for the quantification of modern slavery"
author: "George Linder, Morris Greenberg, Pierre Gardan and Olivier Binette"
date: 'December 14, 2019'
runtime: shiny
output:
  html_document:
    includes:
      before_body: html/header.html
    code_folding: hide
    df_print: paged
    theme: flatly
    toc: true
    number_sections: false
    toc_depth: 2
    toc_float: 
      collapsed: true
---


# 1. Introduction

Forty million people currently live in slavery, according to the International Labour Organization (ILO) [most recent estimate](https://www.ilo.org/global/publications/books/WCMS_575479/lang--en/index.htm). In other words, about every 1 in 220 person worldwide is trapped in a situation of "forced labour, forced marriage, or under another form of exploitation that a person cannot refuse or leave because of threats, violence, coercion, deception and/or abuse of power". **Modern slavery** is an umbrella term which broadly refers to such forms of slavery that are currently happening. 

The scale of the problem has prompted the United Nations to add, in 2015, the eradication of modern slavery as part of its [Sustainable Development Goals framework](https://sustainabledevelopment.un.org/). This calls for integrated action to end modern slavery and for the development of reliable statistical indicators. 

> "Take immediate and effective measures to eradicate forced labour, end modern slavery and human trafficking and secure the prohibition and elimination of the worst forms of child labour, including recruitment and use of child soldiers, and by 2025 end child labour in all its forms." (*Target 8.7 of UN's Sustainable Development goals.*)


<!--
However, statistical methodology to estimate the number of undetected victims of modern slavery is still in development. Victims of modern slavery are often hidden, making reliable estimates of the number of victims particularly challenging to develop.
-->

## 1.1. The case of the UK

The United Kingdom's quantification of modern slavery, together with their 2015 Modern Slavery Act, has drawn considerable attention to such efforts. [Silverman (2014)](https://www.gov.uk/government/publications/modern-slavery-an-application-of-multiple-systems-estimation); [Bales et al. (2015)](https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2015.00824.x) estimated between 10,000 to 13,000 potential victims using a method called **Multiple Systems Estimation** (MSE). The method aggregates data from multiple sources, such as the police and non-governmental organizations, and estimates the proportion of victims that have not been seen by any of them.

However, follow-up research has shown high sensitivity of the estimates to modelling assumptions, with the Walk Free Foundation providing an [about 10-fold larger estimate](https://downloads.globalslaveryindex.org/ephemeral/GSI-2018_FNL_190828_CO_DIGITAL_P-1576254958.pdf) of about 136 000 victims and a recent analysis [(Whitehead et al., 2019)](https://www.tandfonline.com/doi/abs/10.1080/23322705.2019.1660952?journalCode=uhmt20) even suggesting to abandon attempts at measuring this hidden population. It is unclear if the assumptions underlying standard use of MSE were satisfied in Silverman's study and we are concerned by these developments. MSE methods are used in many sensitive human rights applications and their correct usage is critical to reach truth and maintain credibility. 

## 1.2. Our benchmark

We provide some of the tools necessary to investigate the accuracy and stability of MSE estimates in application to the quantification of modern slavery. Precisely, this interactive document (Section 3) allows to:

1. omit or merge sources of information about potential victims and see the effect on the estimates;
2. estimate the coverage of confidence intervals reported by [Silverman (2014)](https://www.gov.uk/government/publications/modern-slavery-an-application-of-multiple-systems-estimation) through a parametric Bootstrap; and
3. simulate data for a range of realistic model and assess the accuracy of MSE in this context. 

We provide an introduction to MSE in Section 2, the tool is embedded in Section 3 and we sumarize our results in Section 4.

# 2. Multiple systems estimation for the quantification of modern slavery





# 3. Benchmark {.tabset}

## Data manipulation

```{r}
shinyAppFile("data_application/shiny.R", options = list(
    width = "100%", height = 1000
  ))
```

## Bootstrap simulation

```{r}
shinyAppFile("simulation/shiny2.R", options = list(
    width = "100%", height = 550
  ))
```

## Generative models

```{r}
shinyAppFile("generative/shiny.R", options = list(
    width = "100%", height = 1000
  ))
```



